### 2022-09-28 paper/results.Rmd
beginning to look like a paper

---
title: "Tables and Figures"
always_allow_html: true
output: 
  officedown::rdocx_document:
    keep_md: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
 echo = FALSE,
 message = FALSE,
 warning = FALSE,
 fig.width = 7,
 fig.height = 4, 
 dpi = 300
)

suppressPackageStartupMessages({
 library(officer)
 library(cardioStatsUSA)
 library(magrittr)
 library(gtsummary)
 library(tidyverse)
 library(survey)
 library(flextable)
 library(plotly)
})

```

__Table 1__ Participant characteristics

```{r, cache=TRUE}

tbl_summary_data <- as_tibble(nhanes_bp) %>% 
 filter(svy_subpop_htn == 1) %>% 
 mutate(nobs = 1) %>% 
 select(
  svy_psu,
  svy_strata,
  svy_weight_mec,
  svy_year,
  nobs,
  starts_with('demo'),
  starts_with('bp'), - bp_cat_meds_excluded,
  starts_with('htn'),
  starts_with('cc')
 )

for(i in names(tbl_summary_data)){
 attr(tbl_summary_data[[i]], 'label') <- nhanes_key$variables[[i]]$label
}

tbl_summary_design <- tbl_summary_data %>% 
 svydesign(ids = ~ svy_psu,
           strata = ~ svy_strata,
           weights = ~ svy_weight_mec,
           data = .,
           nest = TRUE)

tbl_gtsummary <- tbl_summary_design %>% 
 tbl_svysummary(
  by = svy_year,
  include = c(nobs, 
              starts_with("demo"),
              starts_with('bp'), 
              starts_with('htn'),
              starts_with('cc')),
  missing = 'no',
  label = list(nobs ~ "No. of Adults"),
  statistic = list(
   all_categorical() ~ "{p}",
   all_continuous() ~ "{mean} ({sd})",
   nobs ~ "{N_unweighted}"
  )
 ) %>%
 modify_header(all_stat_cols() ~ "**{level}**") %>%
 modify_footnote(all_stat_cols() ~ NA)

as_flex_table(tbl_gtsummary) %>% 
 theme_box() %>% 
 width(width = 0.90) %>% 
 width(j = 1, width = 2) %>% 
 align(align = 'center') %>% 
 align(j = 1, align = 'left')

```

```{r}
block_section(
  prop_section(
    page_size = page_size(orient = "landscape"),
    page_margins = page_mar(bottom = 1/12, top = 1/12, 
                            right = 1/8, left = 1/8),
    type = "continuous"
  )
)
```

__Figure 1__ Prevalence of hypertension by calendar year

```{r}

subset_variables <- list("demo_pregnant")
subset_values <- list(c("No", "Missing"))

nhanes_visualize(outcome = 'htn_accaha', 
                 subset_variables = subset_variables,
                 subset_values = subset_values,
                 years = 'all')[[1]] %>% 
 layout(font = list(size = 40))

```

\newpage

__Figure 2__ Antihypertensive medication use by calendar year among adults with hypertension

```{r}

nhanes_visualize(
 outcome = 'bp_med_use', 
 subset_variables = c(subset_variables, 'htn_accaha'),
 subset_values = c(subset_values, 'Yes'),
 years = 'all'
)[[1]] %>% 
 layout(font = list(size = 40))

```

\newpage

__Figure 3__ Prevalence of resistant hypertension among adults with hypertension taking antihypertensive medication

```{r}

nhanes_visualize(
 outcome = 'htn_resistant_accaha', 
 subset_variables = c(subset_variables, 'htn_accaha', 'bp_med_use'),
 subset_values = c(subset_values, 'Yes', 'Yes'),
 years = 'all'
)[[1]] %>% 
 layout(font = list(size = 40))

```

\newpage

__Figure S1__: Participants included in the current analysis

\newpage

__Figure S2__ Prevalence of hypertension by age group

```{r}

nhanes_visualize(outcome = 'htn_accaha', 
                 exposure = 'demo_age_cat',
                 subset_variables = subset_variables,
                 subset_values = subset_values,
                 years = 'most_recent')[[1]] %>% 
 layout(font = list(size = 40))

```

\newpage

__Figure S3__ Prevalence of hypertension by sex

```{r}

nhanes_visualize(outcome = 'htn_accaha', 
                 exposure = 'demo_gender',
                 subset_variables = subset_variables,
                 subset_values = subset_values,
                 years = 'most_recent')[[1]] %>% 
 layout(font = list(size = 40))

```

\newpage

__Figure S4__ Prevalence of hypertension by race/ethnicity

```{r}

nhanes_visualize(outcome = 'htn_accaha', 
                 exposure = 'demo_race',
                 subset_variables = subset_variables,
                 subset_values = subset_values,
                 years = 'most_recent')[[1]] %>% 
 layout(font = list(size = 40))

```