### 2022-10-01 README.Rmd
jhs retreat

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cardioStatsUSA <a href="https://jhs-hwg.github.io/cardioStatsUSA/"><img src="man/figures/logo.png" align="right" height="139" /></a>

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/jhs-hwg/nhanes-shiny-bp/branch/main/graph/badge.svg)](https://app.codecov.io/gh/jhs-hwg/nhanes-shiny-bp?branch=main)
[![R-CMD-check](https://github.com/jhs-hwg/nhanes-shiny-bp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhs-hwg/nhanes-shiny-bp/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Our goal is to provide a platform for exploration of the National Health and Nutrition Examination Survey (NHANES) data. You can find the online application for exploratory data here: https://bcjaeger.shinyapps.io/nhanesShinyBP/

## Installation

You can install the development version of cardioStatsUSA from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("jhs-hwg/cardioStatsUSA")
```

## Example

The shiny application is hosted on a public server, but you can also run it locally:

```{r, eval=FALSE}

library(cardioStatsUSA)
app_run()

```

## TODO list

- Write vignettes for tests of the app
- Write vignette for medication data
- Replication with AB's group
- Incorporate suppression of unreliable estimates
- Documentation/demos for R functions used in the app
- Definitions/notes for all BP variables (table S1 in manuscript)
- Tutorial videos
- Do we need to use AF weights?

## Validation of results from the shiny application

We have replicated results from prior NHANES publications to verify that routines used to conduct analyses in this app are valid. Based on these tests, we have identified some cases where our app does not exactly match previously published articles, and we note those cases here.

TO BE WRITTEN: notes on tests showing exact match with [Muntner et al](https://jamanetwork.com/journals/jama/fullarticle/2770254)

TO BE WRITTEN: notes on discrepancies with results in [Carey et al.](https://www.ahajournals.org/doi/10.1161/HYPERTENSIONAHA.118.12191)