### 2022-10-21 rmd/nhanes_design_subset.Rmd
big update
added CI options for percentage, tweaked api, fixed many small things

```{r child = 'aaa.Rmd'}

```

# Examples

First we create an NHANES design object

```{r child='make_simple_nhanes_design.Rmd'}

```

Subset calls work as they do in many other `subset` functions:

```{r}

ds_men <- nhanes_design_subset(ds, demo_gender == 'Men')

```

To verify `ds_men` has been subsetted, you can do something like this:

```{r}
table(ds_men$design$variables$demo_gender)
```

and you can also find the subsetted indices here:

```{r}
ds_men$subset_rows
```

Why does the design object save this information? In case you update it in the future and subsetting operations need to be repeated.

## Only one subset is allowed

Due to the way `nhanes_design` objects are handled, it is not a good idea to subset them more than once. In fact, it is currently not allowed. If this becomes problematic, please open an issue on GitHub and let us know!

```{r, error=TRUE}
nhanes_design_subset(ds_men, demo_age_years > 40)
```